---
title: "Nutrients Curacao"
author: "Mike Wit"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
# load packages
# Loading packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, openxlsx, ggmap, 
               sf, leaflet, data.table, cowplot, data.table, knitr,
               pander)

# load data
input <- "C:/Users/mikewit/Documents/SEALINK/Data/Clean_data/final_merged/" 

data <- openxlsx::read.xlsx(paste0(input, "hydrochemistry_curacao.xlsx"))
meta <- openxlsx::read.xlsx(paste0(input, "metadata_2021_2022.xlsx"))

```

## Hydrochemistry and metadata structure

Below the content and structure of the data is given:

```{r cars}
str(data)
kable(head(data))
```

### 2nd fieldwork Oct'22 - Jan'23
Overview sample types

```{r}
data %>%
  filter(year == 2022) %>%
  filter(parameter %in% c("EC", "pH", "NO3", "PO4", "NH4", "HCO3")) %>%
  group_by(subtype) %>%
  ggplot(., aes(x = subtype, y = value, fill = subtype)) +
  geom_boxplot() +
  theme_bw() +
  facet_wrap(facets = "parameter", scales = "free")


```


### Nitrate

Comparing nitrate concentrations in groundwater for the years 1992, 2021 and 2022.

```{r nitrate I, echo=FALSE}
d <- data %>%
  filter(parameter == "NO3",
         sampletype == "groundwater") %>%
  select(year, samplecode, value, limit_symbol)

ggplot(d, aes(x = as.factor(year), y = value, fill = as.factor(year))) +
  geom_boxplot() +
  scale_y_continuous("NO3 [mg/L]") +
  coord_cartesian(ylim = c(0, 200)) +
  scale_x_discrete("") +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle("Nitrate concentrations in groundwater, Curacao")
```
Comparing multiple parameters for 2021 - 2022

```{r}
d <- data %>%
  filter(parameter %in% c("EC", "pH", "NO3", "PO4", "NH4"),
         sampletype == "groundwater") %>%
  select(year, samplecode, parameter, value, limit_symbol)

ggplot(d, aes(x = as.factor(year), y = value)) +
  geom_boxplot() +
  # scale_y_continuous("NO3 [mg/L]") +
  # coord_cartesian(ylim = c(0, 200)) +
  # scale_x_discrete("") +
  theme_bw() +
  # theme(legend.position = "none") +
  # ggtitle("Nitrate concentrations in groundwater, Curacao") +
  facet_wrap(facets = "parameter", scales = "free")


```


For the same wells measured over time: and GW071: Herb garden open well GW024 and GW072: Klein Kwartier well 5Z14(B) GWXXX and GW073: Geert GWXXX and GW074:

```{r table1, echo=FALSE, message = FALSE, results='asis'}
tabl <- "
|  1977  |  1992  |  2020  |  2021  |  2022  | Description |
|--------|--------|--------|--------|--------|-------------|
|        |        |        |        |  GW071 | Herb Garden |
| | | | GW024 | GW072 | Klein Kwartier 5Z14(B) |
| | | | | |
| | | | | |
| | | | | |
"
cat(tabl)
```

```{r}
# only wells with multiple measurements
d <- data %>%
  group_by(putcode) %>%
  mutate(yearly_measurement = n_distinct(year),
         samples = paste(unique(samplecode), collapse = ", ")) %>%
  ungroup() 

d %>%
  filter(parameter == "EC",
         yearly_measurement > 1,
         putcode != "") %>%
  select(putcode, samples, year, value, yearly_measurement) %>%
  distinct() %>%
  pivot_wider(names_from = year,
              values_from = value) %>% view()
  

```

### Phosphate

```{r phosphate I, echo=FALSE}
d <- data %>%
  filter(parameter == "PO4",
         sampletype == "groundwater") %>%
  select(year, samplecode, value, limit_symbol)

ggplot(d, aes(x = as.factor(year), y = value)) +
  geom_boxplot() +
  scale_y_continuous("PO4 [mg/L]") +
  #coord_cartesian(ylim = c(0, 200)) +
  scale_x_discrete("") +
  theme_bw() +
  ggtitle("Phosphate concentrations in groundwater, Curacao")
```
